Seasonality in dominant upper-ocean dynamics at submesoscales

### 2016-05-09 writeup/rocha_etal.tex
Adds appendix

A key aspect of the simulation analyzed here is that it is forced by
the 16 most significant tidal components.
Barotropic tides interact with topography and generate internal
tides that project on mesoscales to submesoscales
\citep{rocha_etal2016}. Hence, tidal forcing fundamentally differentiates our analysis
from recent modeling studies that investigated upper dynamics
and its seasonality \citep{sasaki_etal2014,qiu_etal2014}.

@@

%  this should go in a supplementary material
\appendix
\section{How well does the model captures high-frequency modes?}
 \begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=.9\textwidth]{figs/figA1.pdf}
  \caption{A comparison.}
  \label{figA1}
  \end{center}
\end{figure*}

>>>>

A key aspect of the simulation analyzed here is that it is forced by
the 16 most significant tidal components.
Barotropic tides interact with topography and generate internal
tides that project on mesoscales to submesoscales
\citep{rocha_etal2016}. Hence, tidal forcing fundamentally distinguishes our analysis
from recent modeling studies that investigated upper dynamics
and its seasonality \citep{sasaki_etal2014,qiu_etal2014}.

@@

%  this should go in a supplementary material
\appendix
\section{How well does the model captures high-frequency modes?}
To assess how well the LLC2160 represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.

Figure \ref{figA1} shows KE frequency spectra for the model and mooring data.
At the uppermost common depth (20 m), very low frequencies (periods > 10 days)
and the inertial, diurnal, and semi-diurnal frequencies agree well --- the KESS
 data are more energetic at intermedia frequencies. At 40 m, the model and observations
 have consistent spectra at periods larger than about 6 h. The data from moorings
are more energetic at higher frequencies and tidal harmonics are not observed,
except for the first two harmonics that are apparent in the KESS data at 40 m.
Three reasons for the lack of tidal harmonics in the mooring data are: 1) The high-
frequency variability is dominated by high-frequency noise due to mooring vertical
excursions; 2) The model does not have enough resolution to resolve the full spectrum
of ultraviolet catastrophic wave-wave interactions; 3) The model is not forced at
frequencies higher that 6h.

 \begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=.9\textwidth]{figs/figA1.pdf}
  \caption{A near-surface KE spectra comparison between model and two
  available mooring observations.}
  \label{figA1}
  \end{center}
\end{figure*}
